
# Communicable, Maternal, Neonatal, and Nutritional Disease Burden (CMNN)

```{r echo=FALSE}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here)
#authors: 
#1.  **Partner 1 (Abdullahi)** - Section 1: CMNN burden
#2.  **Partner 2: Garuba** - Section 2: NCD burden
#3.  **Partner 3: Farhad** - Section 3: Overall burden

# Import the communicable diseases data
data_cmnn <- read_csv(here("data/burden-of-disease-cmnn.csv"))


```


```{r}
# Filter the dataset to include only the three countries chosen by your group.
data_cmnn_mini <- data_cmnn %>% 
  filter(Entity %in% c("Afghanistan", "Bangladesh", "Botswana")) %>% 
    rename(DALYs_CMNN = `DALYs (Disability-Adjusted Life Years) - Communicable, maternal, neonatal, and nutritional diseases - Sex: Both - Age: Age-standardized (Rate)`,
           country=Entity)
  
```

## Table of Estimates for CMNN Burden Over Time

```{r}
# Here render a table for the DALY burden over time for the three countries.
# You should pivot the data wider to show each country in a separate column.
data_cmnn_wide <- data_cmnn_mini %>% 
pivot_wider(names_from = Year, values_from=DALYs_CMNN) 
# Use kable() from the knitr package to render the table.
kable(data_cmnn_wide, caption = "DALY Burden Over Time for Selected Countries")
```

## Plot Showing Trends in CMNN Burden Over Time

```{r}
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.
# NOTE: Use your long 'mini' data frame for plotting, not the wide data used for the table. 
ggplot(data=data_cmnn_mini, mapping=aes(x=Year, y= DALYs_CMNN, group=country, color=country, fill=country))+
  geom_line()+
  geom_point()+
  labs(
    x= "Year",
    y="Disability-Adjusted Life Years (DALY)",
    title="Communicable, Maternal, Neonatal, and Nutritional Disease Burden trend",
    subtitle= "Afghanistan, Bangladesh and Botswana, 1990 to 2019",
    caption = "Data Source: daly_report_PARENT"
  )+
theme_minimal()
```

## Summary of CMNN Burden Findings

Provide a brief analysis based on the data presented in the table and chart. Highlight any significant findings or patterns. About 3 sentences.
#Botsana has the highest burden of disease among the three countries. 
#the burden of disease in Botwana has decreased after 2000 onwards 
#Bangladesh had the highest of disease burden in 1990  but it has gradually decreased and it has the lowest burdent of disease
