#group3
#Farhad,Abdullahi,Garuba
# Overall Disease Burden

```{r}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here)

# Import the overall disease data
data_overall <- read_csv(here("data/burden-of-disease-all-causes.csv"))
```

```{r}
# Filter the dataset to include only the three countries chosen by your group.
data_overall_mini <- data_overall %>%
  filter(Entity %in% c("Afghanistan", "Bangladesh", "Botswana"))
```

## Table of Estimates for Overall Disease Burden Over Time

```{r}
# Here render a table for the DALY burden over time for the three countries. 
# You should pivot the data wider to show each country in a separate column.
# Pivot data to show each country in a separate column
data_overall_wide <- data_overall_mini %>%
  select(Entity, Year, DALYs = 4) %>%
  pivot_wider(names_from = Entity, values_from = DALYs)

# Render table using kable
kable(data_overall_wide, caption = "DALY Burden Over Time for Selected Countries")



# Use kable() from the knitr package to render the table.
```

## Plot Showing Trends in Overall Disease Burden Over Time

```{r}
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.
# NOTE: Use your long 'mini' data frame for plotting, not the wide data used for the table. 

daly_plot <- ggplot(data_overall_mini, aes(x = Year, y = `DALYs (Disability-Adjusted Life Years) - All causes - Sex: Both - Age: Age-standardized (Rate)`, color = Entity)) +
  geom_line(size = 1) +
  labs(
    title = "DALY Burden Over Time",
    x = "Year",
    y = "DALY Rate (Age-standardized)",
    color = "Country"
  ) +
  theme_minimal()

# Print the plot
daly_plot

# Save the plot as JPG
ggsave(filename = "daly_burden_over_time.jpg", plot = daly_plot, width = 10, height = 6, dpi = 300)


```

## Summary of Overall Disease Burden Findings

The chart shows a clear decline in DALY rates over time for all three countries. Bangladesh demonstrates the most consistent and steep reduction in disease burden, while Botswana shows a sharp increase until the early 2000s followed by a significant decline. Afghanistan’s burden remains relatively high but gradually decreases over the years.
